### 2024-02-15 paper/paper.qmd
Data section added

---
title: "Racial Force Probe"
author:
  - Monica Sainani
  - Kaavya Kalani
  - Aamishi Avarsekar
thanks: "Code and data are available at: https://github.com/kaavyakalani26/prison_health.git and replication aspects are available at: https://www.socialsciencereproduction.org/reproductions/82b709e2-4aa0-40e8-8abe-195c51ae739b/index"
date: today
date-format: long
abstract: "This research endeavors to scrutinize the well-being of individuals within the prison system, focusing on the prevalence of physical diseases, mental health issues, and mortality rates among incarcerated populations. The prison environment presents unique challenges that can significantly impact the overall health of individuals, making it imperative to investigate the multifaceted dimensions of prisoners' well-being. Utilizing data collected over the past [insert number of years], this study employs a comprehensive analysis to discern patterns, trends, and potential variations in the health outcomes of prisoners. By examining the intersection of physical and mental health alongside mortality rates, the research aims to contribute valuable insights into the current state of prisoners' well-being. This analysis not only serves to inform our understanding of the challenges faced by incarcerated individuals but also provides a foundation for evidence-based policy recommendations aimed at fostering improved conditions within correctional facilities. Through this exploration, we strive to contribute to the ongoing dialogue surrounding the intersection of criminal justice and public health, emphasizing the importance of compassionate and effective strategies for promoting the well-being of those within the prison system."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Preamble ####
# Purpose: Read the cleaned dataset to use it for the analysis
# Author: Kaavya Kalani, Monica Sainani, Aamishi Avarsekar
# Date: 14 February 2024
# Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca, aamishi.avarsekar@mail.utoronto.ca
# License: MIT
# Pre-requisites: run scripts/01-data_cleaning.R

#### Workspace setup ####
# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("here")
# install.packages("ggplot2")
# install.packages("readxl")

library(ggplot2)
library(readxl)
library(tidyverse)
library(knitr)
library(here)
```

# Introduction

Health and Safety of Prisoners Today

Individuals' health and safety in jail are critical aspects of the criminal justice system
that must be thoroughly examined. As we negotiate the complicated world of prisons in contemporary America,
it becomes clear that the physical and emotional health of jailed people is a major problem. This study aims to give a
complete analysis of the health difficulties encountered by inmates today, diving into the areas of physical health,
emotional well-being, death rates, and the subtle distinctions between male and female prisoners.

Physical Health of Prisoners: Unveiling a Disturbing Reality

The first aspect of our investigation is a thorough examination of the physical health of convicts.
Recent discoveries in America portray a troubling picture, demonstrating that a sizable proportion of offenders enter
correctional facilities with a variety of pre-existing physical health issues.
The jail environment, which is marked by overcrowding, restricted access to healthcare, and a lack of preventive
measures, frequently exacerbates these problems(Stevenson, B. (2021)). According to studies, the prison population has a high prevalence of
chronic diseases such as diabetes, hypertension, and infectious diseases, emphasizing the critical need to focus on
the physical well-being of jailed people.

The sad fact is that convicts have far lower health outcomes than the general population.
Access to medical care in prisons remains a recurrent issue, resulting in delayed diagnoses and insufficient treatment.
The cumulative effect of these conditions adds to a cycle of declining physical health among inmates, raising not just
ethical problems but also public health dangers when they are eventually released into society(M. Maruschak, L., & Berzofsky, M. (2015)).

Mental Health: The Silent Crisis Behind Bars

Moving beyond physical health, the second half of our investigation focuses on the sometimes-ignored topic of mental
well-being in the prison system. Mental health difficulties among inmates have reached crisis proportions, with a slew
of obstacles compounding the psychological toll of incarceration.(Edgemon, T. G., & Clay-Warner, J. (2019)). Suicide rates among convicts have become a major
problem, underlining the critical need for mental health services in correctional facilities.

Substance addiction exacerbates the mental health problem since many inmates take drugs to cope with the stresses of
confinement. The absence of comprehensive addiction treatment services in jails reinforces the cycle of substance misuse,
exacerbating mental health issues and impeding successful recovery attempts.(Edgemon, T. G., & Clay-Warner, J. (2019))

The mental health crisis among inmates emphasizes the significance of treating the underlying causes of psychological
anguish in correctional facilities. Initiatives to provide mental health support, provide counselling services, and
establish drug addiction treatment programs are critical to interrupting the cycle of mental health worsening among inmates.

From Suicides to Mortality Rates: Connecting the Dots

The third feature of our investigation focuses on the relationship between suicides and prison death rates.
Suicides in correctional facilities are not isolated instances; they are inextricably linked to larger concerns such as
mental health, poor healthcare, and the harsh reality of incarceration. Beyond suicides, inmates are always at risk of
death from a variety of causes, including murders, infectious infections, and other health difficulties.(Emily Widra, (2017))

Understanding the interconnectedness of these mortality variables is critical for designing holistic measures to improve
prisoner safety and well-being((Emily Widra, (2017)). It demands a comprehensive strategy that tackles both immediate concerns, such as
suicide prevention, as well as the larger structural factors that contribute to higher prison death rates.

Gender Disparities in Prison Health: Unveiling Unique Challenges

In the third portion of our investigation, we examine the subtle disparities in the health experiences of male and female
convicts. While many health issues are common, gender-specific characteristics play an important role in defining the health
outcomes of jailed people. Women inmates, for example, frequently confront particular reproductive health challenges,
as well as an increased risk of sexual assault while in correctional facilities.

Understanding and treating gender differences in prison health is critical for establishing focused therapies that
meet the unique requirements of both male and female inmates. Recognizing the unique issues experienced by each gender
group is an important step toward establishing a more equal and compassionate approach to healthcare in the prison system.

Conclusion: A Call for Holistic Reforms

To summarize, the health and safety of inmates today require immediate attention
and thorough improvements. The complex network of physical health issues, mental health crises,
mortality risks, and gender imbalances inside the jail system needs a comprehensive strategy.
Addressing these concerns necessitates not just urgent actions, such as increased healthcare access and mental health
assistance, but also a larger commitment to rethinking the purpose and consequences of incarceration on individual
well-being(Ghazanfari, H., & Farokhzadian, J. (2022)).This article intends to add to the continuing discussion about prison health by arguing for policies and
practices that prioritize the dignity, safety, and general well-being of people housed in correctional institutions.

# Data

## Source
For this analysis we have used multiple datasets. This paper uses the datasets available in the replication package of the original paper. The author of the original paper has collated a large range of files from multiple sources and we will be using some of the datasets from this collation. The raw datasets obtained from the replication package and cleaned and combined to form the datasets for this paper

## Methodology
For this paper, the datasets were cleaned and analysed using the statistical programming software R [@citeR] along with the help tidyverse [@citeTidyverse], knitr [@citeKnitr], ggplot2 [@citeGgplot], here [@citeHere], zoo [@citeZoo], readxl [@citeRead], writexl [@citeWrite] and janitor [@citeJanitor].

## Datasets

### cleaned_health_data_by_year.csv
This dataset is obtained by cleaning and combining the raw mass-incarceration.xlsx file and msfp0116stt02data.csv.

mass-incarceration.xlsx contains three sheets from which we will be using sheet 1 which contains time series data for incarceration from 1925-2019; each year is associated with the number of imprisoned individuals per 100k people, calculated from the total population and total number of inmates. We clean this file to get the counts of prisoners for 2006-2016.

msfp0116stt02data.csv contains counts for causes of deaths of the prisoners for the years 2001, 2006-2016. For our analysis we will not be considering 2001 but only 2006 to 2016 for the trends as it gives us a continuous period to observe trends over.

Cleaning and combining these lead to our  cleaned_health_data_by_year.csv which contains:
- Year (between 2006 to 2016 inclusive)
- Disease (the cause of death of the prisoner)
- Count (the number of prisoners who died due the corresponding cause in the corresponding year)
- Prisoner_Count (the number of prisoners in that year)
- Proportion (the proportion of prisoners who died due the corresponding cause in the corresponding year)

A person becomes an entry in this dataset if they were a prisoner in any of the prisons in the United States of America between 2006 and 2016 inclusive and died in that year. The reason of the death is recorded among the following causes’categories: Illness, Cancer, Heart Diseases, Liver Diseases, Respiratory disease, AIDS-related, Suicide, Drug/alcohol intoxication, Accident, Homicide, All other reasons or illnesses.

### cleaned_health_data_race.csv and cleaned_health_data_gender.csv
This dataset is obtained by cleaning the raw msfp0116stt09data.csv file.

msfp0116stt09data.csv contains counts for  causes of deaths of the prisoners for the years 2001, 2006-2016 but segmented by demographic factors instead of time, such as gender, race, and age. For our analysis, we will be using the stats segmented by gender and race.

The information segmented by race leads to cleaned_health_data_race.csv which contains:
- x1 (The race of the prisoner)
- Cause (the cause of death of the prisoner)
- Count (the number of prisoners who died due the corresponding cause in the corresponding year)
- Proportion (the proportion of prisoners who died due the corresponding cause in the corresponding year)

A person becomes an entry in this dataset if they were a prisoner in any of the prisons in the United States of America in 2001 or between 2006 and 2016 inclusive and died in that year. The reason of the death is recorded among the following causes’categories: Illness, Cancer, Heart Diseases, Liver Diseases, Respiratory disease, AIDS-related, Suicide, Drug/alcohol intoxication, Accident, Homicide, All other reasons or illnesses. The race of the person is classified as Black, Hispanic, White or Other

The information segmented by race leads to cleaned_health_data_gender.csv which contains:
- x1 (The gender of the prisoner)
- Cause (the cause of death of the prisoner)
- Count (the number of prisoners who died due the corresponding cause in the corresponding year)
- Proportion (the proportion of prisoners who died due the corresponding cause in the corresponding year)

A person becomes an entry in this dataset if they were a prisoner in any of the prisons in the United States of America in 2001 or between 2006 and 2016 inclusive and died in that year. The reason of the death is recorded among the following causes’categories: Illness, Cancer, Heart Diseases, Liver Diseases, Respiratory disease, AIDS-related, Suicide, Drug/alcohol intoxication, Accident, Homicide, All other reasons or illnesses. The gender of the person is either Male or Female in this dataset.

### cleaned_covid19_data.csv
This dataset is obtained by cleaning the raw covidts01.csv file.

covidts01.csv contains the collective number of new cases and case rates per 100,000 among the incarcerated, prison staff, and the general population. The numbers are recorded on a daily basis from April 22, 2020 to the end of August 2020. For our analysis, we just look at the number of new cases per 100,000.

- Cleaning of this file leads to cleaned_covid19_data.csv which contains:
- Date (The date of the new case)
- New_cases_per_100k (The number of new cases recorded on the specific date per 100k)
- Group (The group of people the case belonged to)

A person becomes an entry in this dataset if they were incarcerated or prison staff in any of the prisons or the general population in the United States of America between April 22, 2020 to the end of August 2020 inclusive and contracted Covid-19.

We will now analyse these datasets to uncover trends and draw conclusions.

# Results

```{r}
#| message: false
#| echo: false
#| label: fig-overall-death
#| fig-cap: Causes of deaths among the prisoners

# Load cleaned data from CSV
health_data_by_year_long_prop <- read.csv("data/analysis_data/cleaned_health_data_by_year.csv", header = TRUE, stringsAsFactors = FALSE)

# Create the plot
ggplot(health_data_by_year_long_prop, aes(x = as.numeric(Year), y = Proportion, color = Disease)) +
  geom_line(size = 1.0) +
  labs(title = "Trends of Diseases by Year",
       x = "Year",
       y = "Proportion of cases",
       color = "Disease") +
  theme_minimal()

```

```{r}
#| message: false
#| echo: false
#| label: fig-covid
#| fig-cap: Covid trends in the jail

# Load cleaned data
ts_data <- read.csv("data/analysis_data/cleaned_covid19_data.csv", header = TRUE, stringsAsFactors = FALSE)

# Create the figure using ggplot2 with solid lines
ggplot(ts_data, aes(x = date, y = new_cases_per_100k, color = group)) +
  geom_line(size = 0.8) +  # Adjust the line thickness here (e.g., size = 0.8)
  scale_x_date(date_labels = "%B", date_breaks = "1 month", limits = c(min(ts_data$date), max(ts_data$date))) +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 25)) +
  labs(x = NULL, y = "New cases per 100,000", title = "Seven-Day Moving Average of New Daily COVID-19 Cases per 100,000 Among
Those in Prison, Prison Staff, and in the General Population: May to September
2020") +
  theme_minimal() +
  geom_vline(xintercept = as.numeric(as.Date(c("2020-04-01", "2020-05-01", "2020-06-01", "2020-07-01", "2020-08-01", "2020-09-01"))),
             linetype = "dashed", color = "gray") +
  annotate("text", x = as.Date(c("2020-04-01", "2020-05-01", "2020-06-01", "2020-07-01", "2020-08-01", "2020-09-01")),
           y = 150, label = c("April 2020", "May 2020", "June 2020", "July 2020", "August 2020", "September 2020"),
           vjust = -0.5, hjust = 0, color = "gray") +
  scale_color_manual(values = c("blue", "red", "green"))

```

```{r}
#| message: false
#| echo: false
#| label: fig-death-vs-gender
#| fig-cap: Causes of deaths among the prisoners by gender

# Load cleaned data
health_demographic_data_clean <- read_csv("data/analysis_data/cleaned_health_data_gender.csv")

# Plot the proportions graph using geom_bar with position = "dodge"
ggplot(health_demographic_data_clean, aes(x = x1, y = Proportion, fill = Cause)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +
  labs(title = "Proportions of Causes of Death by Category",
       x = "Gender",
       y = "Proportion") +
  scale_fill_discrete(labels = cause_mapping) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| message: false
#| echo: false
#| label: fig-death-vs-race
#| fig-cap: Causes of deaths among the prisoners by race

# Load cleaned data from CSV
race_health_data_long <- read.csv("data/analysis_data/cleaned_health_data_race.csv", header = TRUE, stringsAsFactors = FALSE)

# Create a bar plot showing proportions with black borders
ggplot(race_health_data_long, aes(x = x1, y = Proportion, fill = Cause)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +  # Add black borders
  labs(title = "Proportions of Causes of Death by Category",
       x = "Category",
       y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Discussion

## Weaknesses and next steps

# Appendix

# References