### 2024-02-15 paper/paper.qmd
initial draft for results section

---
title: "Racial Force Probe"
subtitle: "The study looks on a single scientific argument about whether or not certain police officers discriminate against people based on their race. "
author:
  - Monica Sainani
  - Kaavya Kalani
  - Aamishi Avarsekar
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "This research endeavors to scrutinize the well-being of individuals within the prison system, focusing on the prevalence of physical diseases, mental health issues, and mortality rates among incarcerated populations. The prison environment presents unique challenges that can significantly impact the overall health of individuals, making it imperative to investigate the multifaceted dimensions of prisoners' well-being. Utilizing data collected over the past [insert number of years], this study employs a comprehensive analysis to discern patterns, trends, and potential variations in the health outcomes of prisoners. By examining the intersection of physical and mental health alongside mortality rates, the research aims to contribute valuable insights into the current state of prisoners' well-being. This analysis not only serves to inform our understanding of the challenges faced by incarcerated individuals but also provides a foundation for evidence-based policy recommendations aimed at fostering improved conditions within correctional facilities. Through this exploration, we strive to contribute to the ongoing dialogue surrounding the intersection of criminal justice and public health, emphasizing the importance of compassionate and effective strategies for promoting the well-being of those within the prison system."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
#### Preamble ####
# Purpose: Read the cleaned dataset to use it for the analysis
# Author: Kaavya Kalani, Monica Sainani, Aamishi Avarsekar
# Date: 14 February 2024
# Contact: kaavya.kalani@mail.utoronto.ca, monica.sainani@mail.utoronto.ca, aamishi.avarsekar@mail.utoronto.ca
# License: MIT
# Pre-requisites: run scripts/01-data_cleaning.R

#### Workspace setup ####
# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("here")
# install.packages("ggplot2")
# install.packages("readxl")
# install.packages("gdata")
# install.packages("dplyr")
# install.packages("haven")

library(ggplot2)
library(readxl)
library(tidyverse)
library(knitr)
library(here)
library(gdata)
library(dplyr)
library(haven)
```

# Introduction
Health and Safety of Prisoners Today

Individuals' health and safety in jail are critical aspects of the criminal justice system
that must be thoroughly examined. As we negotiate the complicated world of prisons in contemporary America,
it becomes clear that the physical and emotional health of jailed people is a major problem. This study aims to give a
complete analysis of the health difficulties encountered by inmates today, diving into the areas of physical health,
emotional well-being, death rates, and the subtle distinctions between male and female prisoners.

Physical Health of Prisoners: Unveiling a Disturbing Reality

The first aspect of our investigation is a thorough examination of the physical health of convicts.
Recent discoveries in America portray a troubling picture, demonstrating that a sizable proportion of offenders enter
correctional facilities with a variety of pre-existing physical health issues.
The jail environment, which is marked by overcrowding, restricted access to healthcare, and a lack of preventive
measures, frequently exacerbates these problems(Stevenson, B. (2021)). According to studies, the prison population has a high prevalence of
chronic diseases such as diabetes, hypertension, and infectious diseases, emphasizing the critical need to focus on
the physical well-being of jailed people.

The sad fact is that convicts have far lower health outcomes than the general population.
Access to medical care in prisons remains a recurrent issue, resulting in delayed diagnoses and insufficient treatment.
The cumulative effect of these conditions adds to a cycle of declining physical health among inmates, raising not just
ethical problems but also public health dangers when they are eventually released into society(M. Maruschak, L., & Berzofsky, M. (2015)).

Mental Health: The Silent Crisis Behind Bars

Moving beyond physical health, the second half of our investigation focuses on the sometimes-ignored topic of mental
well-being in the prison system. Mental health difficulties among inmates have reached crisis proportions, with a slew
of obstacles compounding the psychological toll of incarceration.(Edgemon, T. G., & Clay-Warner, J. (2019)). Suicide rates among convicts have become a major
problem, underlining the critical need for mental health services in correctional facilities.

Substance addiction exacerbates the mental health problem since many inmates take drugs to cope with the stresses of
confinement. The absence of comprehensive addiction treatment services in jails reinforces the cycle of substance misuse,
exacerbating mental health issues and impeding successful recovery attempts.(Edgemon, T. G., & Clay-Warner, J. (2019))

The mental health crisis among inmates emphasizes the significance of treating the underlying causes of psychological
anguish in correctional facilities. Initiatives to provide mental health support, provide counselling services, and
establish drug addiction treatment programs are critical to interrupting the cycle of mental health worsening among inmates.

From Suicides to Mortality Rates: Connecting the Dots

The third feature of our investigation focuses on the relationship between suicides and prison death rates.
Suicides in correctional facilities are not isolated instances; they are inextricably linked to larger concerns such as
mental health, poor healthcare, and the harsh reality of incarceration. Beyond suicides, inmates are always at risk of
death from a variety of causes, including murders, infectious infections, and other health difficulties.(Emily Widra, (2017))

Understanding the interconnectedness of these mortality variables is critical for designing holistic measures to improve
prisoner safety and well-being((Emily Widra, (2017)). It demands a comprehensive strategy that tackles both immediate concerns, such as
suicide prevention, as well as the larger structural factors that contribute to higher prison death rates.

Gender Disparities in Prison Health: Unveiling Unique Challenges

In the third portion of our investigation, we examine the subtle disparities in the health experiences of male and female
convicts. While many health issues are common, gender-specific characteristics play an important role in defining the health
outcomes of jailed people. Women inmates, for example, frequently confront particular reproductive health challenges,
as well as an increased risk of sexual assault while in correctional facilities.

Understanding and treating gender differences in prison health is critical for establishing focused therapies that
meet the unique requirements of both male and female inmates. Recognizing the unique issues experienced by each gender
group is an important step toward establishing a more equal and compassionate approach to healthcare in the prison system.

Conclusion: A Call for Holistic Reforms

To summarize, the health and safety of inmates today require immediate attention
and thorough improvements. The complex network of physical health issues, mental health crises,
mortality risks, and gender imbalances inside the jail system needs a comprehensive strategy.
Addressing these concerns necessitates not just urgent actions, such as increased healthcare access and mental health
assistance, but also a larger commitment to rethinking the purpose and consequences of incarceration on individual
well-being(Ghazanfari, H., & Farokhzadian, J. (2022)).This article intends to add to the continuing discussion about prison health by arguing for policies and
practices that prioritize the dignity, safety, and general well-being of people housed in correctional institutions.

# Data {#sec-data}

# Model

# Results

Based on the original paper’s findings, we decided to further investigate the health and general well-being of prisoners and trends of diseases within prison, causes of death and also the trends of COVID-19 across a 7-day period.

```{r}
#| message: false
#| echo: false
#| label: fig-overall-death
#| fig-cap: Causes of deaths among the prisoners

# Load cleaned data from CSV
health_data_by_year_long_prop <- read.csv(file=here("data/analysis_data/cleaned_health_data_by_year.csv"), header = TRUE, stringsAsFactors = FALSE)

# Create the plot
ggplot(health_data_by_year_long_prop, aes(x = as.numeric(Year), y = Proportion, color = Disease)) +
  geom_line(size = 1.0) +
  labs(title = "Trends of Diseases by Year",
       x = "Year",
       y = "Proportion of cases",
       color = "Disease") +
  theme_minimal()

```
Figure 1 compares shows the cause of death among prisoners due to diseases. In order to accurately investigate, we adjusted the number of deaths caused by a particular disease by the total number of prisoners recorded for that year. This accounts for the differences in the number of deaths recorded between each cause.

By observing the trends in Figure 1, we can see that the deaths due to cancer and heart diseases is significantly higher compared with other causes. The trends for death due to homicide saw a slight decrease between the later years which is opposite to the rate of deaths caused by suicide. While deaths caused by respiratory diseases remain fairly consistent, there is a gradual increase throughout the years. We can also observe that there is a steady decline in the deaths caused by AIDS. Deaths caused by accident remain fairly uncommon with only a very slight increase in 2012. However, the deaths caused by heart disease remained fairly high yet consistent but saw a steep incline after 2012. The deaths caused by cancer has the highest rate. To take a closer look at this, we can investigate the age-group of the prisoners.

```{r}
#| message: false
#| echo: false
#| label: fig-covid
#| fig-cap: Covid trends in the jail

# Load cleaned data
ts_data <- read.csv(file=here("data/analysis_data/cleaned_covid19_data.csv"), header = TRUE, stringsAsFactors = FALSE)
# Load cleaned data
ts_data <- read.csv(file=here("data/analysis_data/cleaned_covid19_data.csv"), header = TRUE, stringsAsFactors = FALSE)

#Convert date column to Date type
ts_data$date <- as.Date(ts_data$date, format = "%Y-%m-%d")

# Create the figure using ggplot2 with solid lines
ggplot(ts_data, aes(x = date, y = new_cases_per_100k, color = group)) +
  geom_line(size = 0.8) +  # Adjust the line thickness here (e.g., size = 0.8)
  scale_x_date(date_labels = "%B", date_breaks = "1 month", limits = c(min(ts_data$date), max(ts_data$date))) +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 25)) +
  labs(x = NULL, y = "New cases per 100,000", title = "Seven-Day Moving Average of New Daily COVID-19 Cases per 100,000 Among
#Those in Prison, Prison Staff, and in the General Population: May to September
#2020") +
  theme_minimal() +
  geom_vline(xintercept = as.numeric(as.Date(c("2020-04-01", "2020-05-01", "2020-06-01", "2020-07-01", "2020-08-01", "2020-09-01"))),
             linetype = "dashed", color = "gray") +
  annotate("text", x = as.Date(c("2020-04-01", "2020-05-01", "2020-06-01", "2020-07-01", "2020-08-01", "2020-09-01")),
           y = 150, label = c("April 2020", "May 2020", "June 2020", "July 2020", "August 2020", "September 2020"),
          vjust = -0.5, hjust = 0, color = "gray") +
  scale_color_manual(values = c("blue", "red", "green"))

```
result 2

```{r}
#| message: false
#| echo: false
#| label: fig-death-vs-gender
#| fig-cap: Causes of deaths among the prisoners by gender

cause_mapping <- c(
  "cancer" = "Cancer",
  "heart_disease" = "Heart Disease",
  "liver_disease" = "Liver Disease",
  "respiratory_disease" = "Respiratory Disease",
  "aids_related" = "Aid related",
  "all_other_illnesses" = "All other illnesses",
  "suicide" = "Suicide",
  "drug_alcohol_intoxication" = "Drug/Alcohol Intoxication",
  "accident" = "Accident",
  "homicide_d" = "Homicide"
)

# Load cleaned data
health_demographic_data_clean <- read_csv(file=here("data/analysis_data/cleaned_health_data_gender.csv"))

# Plot the proportions graph using geom_bar with position = "dodge"
ggplot(health_demographic_data_clean, aes(x = x1, y = Proportion, fill = Cause)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +
  labs(title = "Proportions of Causes of Death by Category",
       x = "Gender",
       y = "Proportion") +
  scale_fill_discrete(labels = cause_mapping) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Figure 3 compares the cause of death due to diseases between female and male prisoners. The data in the graph has also been adjusted by proportion of female and male prisoners to account for the low counts of reported diseases in females as well as their population. Both graphs have a similar general unimodal shape. Both sexes have the highest deaths caused by cancer and heart diseases, with males with a slightly higher rate than females. However, the least deaths in females are caused by homicide where its rate is almost negligible. While homicide in males is significantly less compared to other death causes, it is still significantly higher compared to its female counter part. On the other hand, the least deaths in males occurs due to accident. Liver disease and respiratory disease are almost comparable in females whereas for males, there is a steep decline in deaths caused by respiratory diseases in comparison to deaths caused by liver disease.

```{r}
#| message: false
#| echo: false
#| label: fig-death-vs-race
#| fig-cap: Causes of deaths among the prisoners by race

# Load cleaned data from CSV
race_health_data_long <- read.csv(file=here("data/analysis_data/cleaned_health_data_race.csv"), header = TRUE, stringsAsFactors = FALSE)

# Create a bar plot showing proportions with black borders
ggplot(race_health_data_long, aes(x = x1, y = Proportion, fill = Cause)) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +  # Add black borders
  labs(title = "Proportions of Causes of Death by Category",
       x = "Category",
       y = "Proportion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Figure 4 compares the causes of deaths in prison with respect to the different races of the prisoners. It can be observed that the cause of deaths are unimodular across all races but they have different peaks in their modes. Amongst Black prisoners, heart diseases have the highest death rate while in Hispanic, White and other races, cancer has the highest death rate. Amongst the identified races, Hispanic prisoners have the highest suicide and homicide rates and also respiratory diseases. It can also be observed that Black prisoners have the highest rate of aids-related deaths, and White prisoners have the highest rate of deaths caused by heart diseases. While deaths caused by drug/alcohol intoxication remains comparatively low across all races, it is the highest for Hispanic prisoners. Amongst other races, it can be observed that suicide is the highest recorded reason for death which could open way to an analyses for racism within prisons as well.

## Discussion

## Weaknesses and next steps

# Appendix {-}

# Additional data details

# References